Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-05-17 causal-bandit/nips_2016/tech.tex
first draft
corrects definition of m(q)
adds statement and proof of the relationship between m(eta) and m(q)

>>>>

\subsection{Relationship between $m(\eta)$ and $m(\boldsymbol{q})$}

\begin{proposition}\label{pro:m-equivelence} In the parallel bandit setting,
$m(\eta^*) \leq 2m(\boldsymbol{q})$.
\end{proposition}

\begin{proof}

Recall that in the parallel bandit setting,

\eq{
\mathcal{A} = \set{do()} \cup \set{ do(X_i = j) \colon 1 \leq i \leq N \text{ and } j \in \set{0,1}}
}

Let:

\eq {
\eta_a = \ind{\P{X_i = j} < \frac{1}{m(\boldsymbol{q})}}\frac{1}{2m(\boldsymbol{q})} \text { for } a \in do(X_i = j)
}

Let $D =\sum_{a\in do(X_i=j)}\eta_a$. From the definition of $m(\boldsymbol{q})$, 
\eq {
\sum_{a\in do(X_i=j)} \ind{\P{X_i = j} < \frac{1}{m(\boldsymbol{q})}} \leq m(\boldsymbol{q}) \implies D \leq \frac{1}{2}
}

Let $\eta_a = \frac{1}{2} + (1-D)$ for $a = do()$ such that $\sum_{a \in \calA}\eta_a = 1$

Recall that,

\eq{
m(\eta) &
= \max_a \EEa\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]
}

We now show that our choice of $\eta$ ensures $\EEa\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] \leq 2m(\boldsymbol{q})$ for all actions $a$.

For the actions $a: \eta_a > 0$, ie $do()$ and $do(X_i = j):\P{X_i=j}<\frac{1}{m(\boldsymbol{q})}$,
\eq{
\EEa\left[\frac{\Pn{a}{X_1...X_N}}{\sum_{b}\eta_b\Pn{b}{X_1...X_N}}\right] \leq \EEa\left[\frac{\Pn{a}{X_1...X_N}}{\eta_a\Pn{a}{X_1...X_N}}\right] = \EEa\left[\frac{1}{\eta_a}\right] \leq 2m(\boldsymbol{q})
}

For the actions $a :\eta_a = 0$, ie $do(X_i=j):\P{X_i=j}\geq\frac{1}{m(\boldsymbol{q})}$,
\eq{
\EEa\left[\frac{\Pn{a}{X_1...X_N}}{\sum_{b}\eta_b\Pn{b}{X_1...X_N}}\right] \leq & \EEa\left[\frac{\ind{X_i=j}\prod_{k\neq i}\P{X_k}}{(1/2+D)\prod_k \P{X_k}}\right] \\=& \EEa\left[\frac{\ind{X_i=j}}{(1/2+D)\P{X_i = j}}\right]
\leq  \EEa\left[\frac{\ind{X_i=j}}{(1/2)(1/m(\boldsymbol{q}))}\right] \leq 2m(\boldsymbol{q})
}

\end{proof}
\
%in the parallel bandit setting the $m(\eta)$ given in this section approximately coincides with the $m(\vec{q})$ in Eq.\ \ref{eq:m-simple}.
%Recall in that setting that 
%\eq{
%\actions = \set{do()} \cup \set{do(X_i = j) : 1 \leq i \leq N \text{ and } j \in \set{0,1}}\,.
%}
%For $a = do()$ choose $\eta_a = 1/2$. 
%Let $m = m(\vec{q})$ and note that, by definition, there are at most $m$ pairs $(i,j)$ such that $\P{X_i = j} \leq 1/m$.
%Thus, for $a = do(X_i = j)$ letting $\eta_a \propto \ind{\P{X_i = j} \leq 1/m} / (2m)$ guarantees $\eta_a \ge \frac{1}{2m}$ when $\eta_a \ne 0$.
%It is then easy to check that $m(\eta) \leq 2m$ using an argument like that for Proposition~\ref{pro:m-bound}.
